---
title: "explore"
output: html_document
---

```{r libraries}
library(dplyr)
library(ggplot2)
library(zoo)
library(Metrics)
```

Create dataframe for:
*gamelogs
*players
*games
```{r load_data}
setwd('~/fantasy_sports/nba/raw_data')
temp <- list.files(pattern = '*.csv')
data_list <- lapply(temp, read.csv, stringsAsFactors = F)
data <- do.call('rbind', data_list)
rm(temp, data_list)
setwd('~/fantasy_sports/nba/R_code')

names <- data %>%
  dplyr::select(
    player_id,
    player_name
  ) %>%
  unique()

games <- data %>%
  dplyr::select(
    season_id,
    game_id,
    team,
    opp
  ) %>%
  unique()
```

```{r ppm history}
ppm <- data %>% 
  dplyr::group_by(
    season_id,
    player_id
  ) %>% 
  dplyr::summarise(
    ppm = mean(dk_pts) / mean(min)
  ) %>% 
  dplyr::left_join(
    y = names,
    by = 'player_id'
  ) %>%
  dplyr::arrange(
    -ppm
  )
```

```{r cats_def}
catsdef_game <- data %>%
  dplyr::group_by(
    season_id,
    opp,
    game_date,
    game_id
  ) %>%
  dplyr::summarise_each(
    funs(sum),
    pts,
    fg3m,
    reb,
    ast,
    stl,
    blk,
    tov
  )

catsdef_season <- catsdef_game %>%
  dplyr::group_by(
    season_id,
    opp
  ) %>%
    dplyr::summarise_each(
    funs(mean),
    pts,
    fg3m,
    reb,
    ast,
    stl,
    blk,
    tov
  )

league_avg <- catsdef_season %>%
  dplyr::group_by(
    season_id
  ) %>%
    dplyr::summarise_each(
    funs(mean),
    pts,
    fg3m,
    reb,
    ast,
    stl,
    blk,
    tov
  )

names(league_avg)[2:length(league_avg)] <- paste0('lgavg_', names(league_avg)[2:length(league_avg)])

catsdef_season_adj <- catsdef_season %>%
  dplyr::left_join(
    y = league_avg,
    by = 'season_id'
  ) %>%
  dplyr::mutate(
    adj_pts = pts / lgavg_pts,
    adj_fg3m = fg3m / lgavg_fg3m,
    adj_reb = reb / lgavg_reb,
    adj_ast = ast / lgavg_ast,
    adj_stl = stl / lgavg_stl,
    adj_blk = blk / lgavg_blk,
    adj_tov = tov / lgavg_tov
  ) %>%
  dplyr::select(
    season_id,
    opp,
    adj_pts,
    adj_fg3m,
    adj_reb,
    adj_ast,
    adj_stl,
    adj_blk,
    adj_tov
  )
```

How consistent is a team's defensive performance over the course of the season when using different rolling window sizes?

```{r window function}
roll_n <- function(data, size) {
  rolled <- apply(
    data[, 5:11],
    2,
    zoo::rollmean,
    k = size
  ) %>%
    as.data.frame()
  
  df <- apply(rolled, 2, round, digits = 2) %>% as.data.frame()
  df$opp <- data$opp[1]
  df$season_id <- data$season_id[1]
  df$g <- size:(nrow(df)+(size-1))
  
  means <- apply(data[5:11], 2, function(x) round(mean(x), digits = 2))
  stats <- names(data)[5:11]
  for (stat in stats) {
    df[, paste0(stat, '_avg')] <- means[match(stat, stats)]
  }
  
  return(df)
}
```

```{r def plotter}
plotter <- function(df, stat) {
  avg <- df[1, paste0(stat, '_avg')] %>% as.numeric()
  g <- ggplot(df) +
    geom_point(
      aes_string(
        x = names(df)[match('g',names(df))],
        y = names(df)[match(stat, names(df))]
      )
    ) +
    geom_line(
      aes_string(
        x = names(df)[match('g',names(df))],
        y = names(df)[match(stat, names(df))]
      )
    ) +
    geom_hline(
      yintercept = avg
    ) +
    ggtitle(
      paste(df$season_id[1], df$opp[1], stat, sep = '_')
    ) +
    xlab('game')
    ylab(stat)
  
  print(g)
}
```

```{r rmse table}
rmse_table <- . %>%
  dplyr::group_by(
    season_id
  ) %>%
  dplyr::summarise(
    pts_rmse = Metrics::rmse(pts, pts_avg),
    fg3m_rmse = Metrics::rmse(fg3m, fg3m_avg),
    reb_rmse = Metrics::rmse(reb, reb_avg),
    ast_rmse = Metrics::rmse(ast, ast_avg),
    stl_rmse = Metrics::rmse(stl, stl_avg),
    blk_rmse = Metrics::rmse(stl, blk_avg),
    tov_rmse = Metrics::rmse(tov, tov_avg)
  ) %>% 
  dplyr::summarise_each(
    funs(mean),
    -season_id
  )
```

```{r create def w/ window}
season_opp_grouper <- . %>%
  dplyr::group_by(
    season_id,
    opp
  )

def5 <- catsdef_game %>%
  season_opp_grouper() %>%
  dplyr::do(roll_n(., 5))

def7 <- catsdef_game %>%
  season_opp_grouper() %>%
  dplyr::do(roll_n(., 7))

def10 <- catsdef_game %>%
  season_opp_grouper() %>%
  dplyr::do(roll_n(., 10))
```

```{r view summary}
View(
  def5 %>%
    rmse_table()
)

View(
  def7 %>%
    rmse_table()
)

View(
  def10 %>%
    rmse_table()
)
```

Assume a player's season average is player's "true" ability level. See if adjusting by opponents rates come close to predicting next game production.

```{r player averages}
rounded_avg <- function(x) {
  return(round(mean(x), digits = 1))
}

player_avg <- data %>%
  dplyr::select(
    season_id,
    player_id,
    min,
    pts,
    fg3m,
    reb,
    ast,
    stl,
    blk,
    tov,
    dub_dub,
    trip_dub,
    dk_pts
  ) %>%
  dplyr::group_by(
    season_id,
    player_id
  ) %>%
  dplyr::summarise_each(
    .,
    funs = funs(rounded_avg)
  ) %>%
  dplyr::mutate(
    avg_dkpts = dk_pts
  ) %>%
  dplyr::select(
    -dk_pts
  )
```

```{r create dataset}
adjusted <- data %>%
  dplyr::select(
    season_id,
    player_id,
    player_name,
    game_id,
    team,
    opp,
    dk_pts
  ) %>%
  dplyr::mutate(
    actual_pts = dk_pts
  ) %>%
  dplyr::left_join(
    y = player_avg,
    by = c('season_id', 'player_id')
  ) %>%
  dplyr::left_join(
    y = catsdef_season_adj,
    by = c('season_id', 'opp')
  ) %>%
  dplyr::mutate(
    pred_pts = (
      pts * adj_pts + fg3m * adj_fg3m + 1.25 * ast * adj_ast + 
        .5 * reb * adj_reb + 2 * stl * adj_stl + 2 * blk * adj_blk - 
        .5 * tov * adj_tov + 1.5 * dub_dub + 3 * trip_dub
    )
  ) %>%
  dplyr::select(
    season_id,
    player_id,
    player_name,
    game_id,
    team,
    opp,
    dk_pts,
    avg_dkpts,
    pred_pts
  )
```

```{r adjusted plot}
avg_plot <- ggplot(adjusted, aes(x = pred_pts, y = dk_pts)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, color = 'red')
```

RMSE using season average as "true" ability and adjusting by opponent's defense:
`r mean(sqrt((adjusted$dk_pts - adjusted$pred_pts)^2))` fantasy points.

Next, break down stats to per minute level and adjust.